---
title: "Datasets for Ivey Business Statistics"
output:
  html_document:
    theme: united
    toc: true
    toc_float: true
css: laurens_styles.css
---

```{r, child="_Global-Options.Rmd"}
```
<br>  
<br>  
<br>  


Below you will find links to the case data files.  

* You can link to .TXT files directly using the url  
* You can download .CSV files to your computer  


<br>  
<br>  
<br>  

## How to import data into R
Access the TXT file directly by url:
```{r, eval=FALSE}
data1 <- read.table(url("https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/data.txt"), 
                        header = TRUE)
```

To download the file, right click on CSV and choose "Save link as... "
Access the CSV file saved onto your computer:
```{r, eval=FALSE}
 data2 <- read.csv("C:/Users/Lauren/FavouriteClass/data.csv", 
                        header = TRUE)
```

<br>  
<br>  
<br>  

## Datasets for Business Statistics
### Cyberville Families
* [Data dictionary](Cyberville-Description.html)  
* Dataset:  [TXT](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/families.txt), [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/families.csv)

<br>  
<br>  

### Medical Expenditure Panel Survey data
* [Data dictionary](MEPS-Description.html)  
* Dataset:  [TXT](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/mepsData.txt), [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/mepsData.csv)

<br>  
<br>  

### Occupational Prestige data
* Data dictionary  
* Dataset:  [TXT](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/PrestigeData.txt), [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/PrestigeData.csv)  

<br>  
<br>  

### Roulette Wheel
* Exhibit 1: [TXT](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/RouletteData_Ex1.txt), [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/RouletteData_Ex1.csv)
* Exhibit 2: [TXT](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/RouletteData_Ex2.txt), [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/RouletteData_Ex2.csv)

<br>  
<br>  

### PGA Golf
* Dataset: [TXT](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/PGAGolfData.txt), [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/PGAGolfData.csv)  

<br>  
<br>  

### Ian Steele
* Dataset: [TXT](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/IanSteeleData.txt), [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/IanSteeleData.csv)  
   
<br>  
<br>  

### Bikeshare
* Dataset: [TXT](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/BikeshareData.txt), [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/BikeshareData.csv)  
   
<br>  
<br>  

### Vungle
* Dataset: [TXT](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/VungleData.txt), [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/VungleData.csv)  
  
<br>  
<br>  

### Pilgrim Bank
* Dataset: [TXT](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/PilgrimBankData.txt), [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/PilgrimBankData.csv)  
   
<br>  
<br>  

### Progressive Insurance
* Dataset: [TXT](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/ProgressiveInsuranceData.txt), [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/ProgressiveInsuranceData.csv)  
  
<br>  
<br>  

### Lansink
* Dataset: [TXT](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/LansinkData.txt), [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/LansinkData.csv)  
  
<br>  
<br>  

### Vanderbilt
* Dataset: [TXT](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/VanderbiltData.txt), [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/VanderbiltData.csv)  
  
<br>  
<br>  

### BigEast Bank
* Dataset: [TXT](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/BigEastData.txt), [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/BigEastData.csv)  

<br>  
<br>  

### Pennsylvania Dominion
* Dataset: [TXT](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/MortgageData.txt), [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/MortgageData.csv)  

<br>  
<br>  

### Happiness
* Dataset: [TXT](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/GSS_HappyData.txt), [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/GSS_HappyData.csv)  
  
<br>  
<br>  

### Select Collections
* Training Data:  [TXT](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/SC_TrainData.txt)   [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/SC_TrainData.csv)  
* Testing Data: [TXT](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/SC_TestData.txt), [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/SC_TestData.csv)  
* Testing Data Paid Amount: [TXT](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/SC_TestPaid.txt), [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/SC_TestPaid.csv)  

 
