---
title: "Datasets for Ivey Business Statistics"
output:
  html_document:
    theme: united
    toc: true
    toc_float: true
css: laurens_styles.css
---

```{r, child="_Global-Options.Rmd"}
```
***
***

## Introduction

Below you will find links to the case data files.  

* You can link to .TXT files directly using the url  
* You can download .CSV files to your computer  

***

## How to import data into R
Access the CSV file directly by url:
```{r, eval=FALSE}
data1 <- read.csv(url("https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/data.csv"), 
                        header = TRUE)
```

To download the file, right click on CSV and choose "Save link as... "
Access the CSV file saved onto your computer:
```{r, eval=FALSE}
 data2 <- read.csv("C:/Users/Lauren/FavouriteClass/data.csv", 
                        header = TRUE)
```

*** 

## Datasets for Business Statistics
### Cyberville Families
* [Data dictionary](Cyberville-Description.html)  
* Dataset:   [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/families.csv)

***

### Medical Expenditure Panel Survey data
* [Data dictionary](MEPS-Description.html)  
* Dataset:   [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/mepsData.csv)

***

### Occupational Prestige data
* [Data dictionary](Prestige-Data-Description.html) 
* Dataset:  [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/PrestigeData.csv)  

***  

### Happiness
* [Data dictionary](Happy-Data-Description.html) 
* Dataset:  [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/GSS_HappyData.csv)  

*** 

### Roulette Wheel
* Exhibit 1:  [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/RouletteData_Ex1.csv)
* Exhibit 2: [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/RouletteData_Ex2.csv)

*** 

### PGA Golf
* Dataset:  [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/PGAGolfData.csv)  

***

### Ian Steele
* Dataset:  [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/IanSteeleData.csv)  
   
***

### Bikeshare
* Dataset:  [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/BikeshareData.csv)  
   
***

### Vungle
* Dataset: [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/VungleData.csv)  
  
***

### Pilgrim Bank
* Dataset:  [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/PilgrimBankData.csv)  
   
***

### Progressive Insurance
* Dataset: [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/ProgressiveInsuranceData.csv)  
  
***

### Lansink
* Dataset:  [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/LansinkData.csv)  
  
***

### Vanderbilt
* Dataset:  [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/VanderbiltData.csv)  
  
*** 

### Direct Marketing
* Dataset:  [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/SubscribeBank.csv)  
* [Data dictionary](Subscription-Description.html) 

*** 

### BigEast Bank
* Dataset:  [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/BigEastData.csv)  

*** 

### Pennsylvania Dominion
* Dataset:  [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/MortgageData.csv)  

  
***

### Select Collections
* Training Data:     [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/SC_TrainData.csv)  
* Testing Data:  [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/SC_TestData.csv)  
* Testing Data Paid Amount: [CSV](https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/SC_TestPaid.csv)  

***
***
