---
title: "Data Manipulation In R"
output:
  html_document:
    theme: united
    toc: true
    toc_float: true
css: laurens_styles.css
---

```{r, child="_Global-Options.Rmd"}
```
***
***

## Introduction
We will use the MEPS dataset for the purposes of this tutorial. The MEPS dataset is a population dataset of medical expenditure data in the United States.
```{r}
data <- read.table(url("https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/mepsData.txt"), header = TRUE)
```

***

## Making new rows and columns

We can use `rbind` and `cbind` to add new rows and columns respectively. They must have the same structure in order to work.
```{r}
sample_row = data[1337, ] # Picking a random row for the example

data_subset = rbind(data, sample_row)
nrow(data)
nrow(data_subset)
```

An easier way to make new columns when working with data.frames is by using the `$`.
```{r}
data2 = data # Keep our original dataset untouched.
data2$newColumn = 1 # Fill the column entirely with 1's
summary(data2$newColumn)
```


***

## Taking Slices of Data
Often you will find that you require only portions of a dataset.

The most basic method is to directly pass the row or column number or name.

```{r}
data_subset = data[ , c("Person_ID", "Age", "Sex")] # Keep these columns
summary(data_subset)
```


```{r}
data_subset = data[1:10000, ] # Keeping the first 10000 rows
summary(data_subset)
```
We will want to do more complex slices as our data may contain outliers, or you are investigating specific individuals.

Let's take a look at some useful functions.

### Subset

```{r}
hist(data$FamIncome_Continuous)
```

As you can see, the majority of data points lie below \$300,000. Let's take a clearer look at those individuals by taking advantage of the `subset` function.
```{r}
data_subset = subset(data, FamIncome_Continuous<300000)
hist(data_subset$FamIncome_Continuous)
```

The dataset now only contains individuals below our target income. We can investigate the impact this has on the entire dataset below.
```{r, collapse=TRUE}
summary(data$FamIncome_Continuous)
summary(data_subset$FamIncome_Continuous)
```

> The `subset` function can be used on vectors, matrices and data frames!

### Which
For even more complex slices of data, the `which` function comes in handy. It works a little differently...the `which` function provides indices where our logic is true, so we have to pass the `which` function as an input for the rows to keep.

Let's investigate individuals with Normal BMI classification (18.5-25) who have diabetes.
```{r}
# Will provide a vector telling us which indices are TRUE
index = which(data$HasDiabetes==1 & #Combine statements with &
              data$BMI>18.5 &
              data$BMI<25)

head(index) #See the first six elements of index

data_subset = data[index, ] # We leave the column input blank to include all columns
```

```{r, collapse=TRUE}
summary(data_subset$HasDiabetes)
summary(data_subset$BMI)
```

### Removing NA's
We may be investigating the impact of hypertension. As you can see below, the `HasHypertension` column has some missing entries. 
```{r}
summary(data$HasHypertension)
```

The `is.na` function tests whether a variable contains NA.
```{r, collapse=TRUE}
is.na(1)
is.na(NA)
is.na(c(1,2,3,NA))
```

We can use the `is.na` function combined with either of our methods above.
```{r}
data_subset = subset(data, is.na(HasHypertension)==0)
# or
index = which(is.na(data$HasHypertension)==0)
data_subset = data[index, ]
```

The data now contains only rows where hypertension is not missing.
```{r}
summary(data$HasHypertension)
summary(data_subset$HasHypertension)  
```

#### Replacing Missing Data
We can take this one step further. Rather than simply removing missing data, we can replace is with the population average. Notice BMI has many missing entries.
```{r}
summary(data$BMI)
```
Let's replace it with the average BMI from the remaining population.
```{r}
# Keeping our original dataset untouched
data2 = data

# First find the indices where BMI is missing
index = which(is.na(data2$BMI)==1)

# Replace the BMI at those indices
data2[index, "BMI"] = mean(data2$BMI, na.rm = TRUE)

summary(data2$BMI)
```

> We can use any method to replace the missing data, try using the `median` and `mode` functions on your own!

### Merge

If we try to use `rbind` to combine datasets we may get some unwanted behaviour.
```{r, collapse=TRUE}
data_subset_diab = subset(data, HasDiabetes==1)
data_subset_hyper = subset(data, HasHypertension==1)

data_subset = rbind(data_subset_diab, data_subset_hyper) # Use rbind to directly combine the tables by row

nrow(unique(data_subset)) # Number of unique observations
nrow(data_subset) # Total Observations
```
As you can see, there are more total observations than unique observations. We are double counting some observations as there are people with both diabetes and hypertension. This is where the `merge` function comes in handy. 

```{r}
data_subset = merge(data_subset_diab, data_subset_hyper, by = "Person_ID", all = TRUE, sort = TRUE)

nrow(data_subset)
```

The `all` parameter tells the function to keep any observation that does not have a matching ID, achieving the behaviour we desire. 

> Remember to check out the help page for any function (using `help(merge)`), they have even more useful information!

***

## Converting Continuous to Categorical (Factors)
Sometimes we may want to investigate our variables as categories. We can convert columns in our table easily. Let's convert our continuous BMI variable into commonly used classifications. We will use the dataset where we replaced missing values with mean BMI.
```{r}
BMICategories = c("Underweight", "Normal Weight", "Overweight", "Obese")

# Initializing column with zeros
data2$BMICategory = 0
```

We must first initialize our categories with simple integers, we will use 1-4 to represent our four categories. The logic within the square brackets tells our data.frame to update only the rows where the statement is true.

```{r}
# Underweight
data2$BMICategory[ data2$BMI<18.5 ] = 1
# Normal Weight
data2$BMICategory[ data2$BMI>=18.5 & data2$BMI<25 ] = 2
# Overweight
data2$BMICategory[ data2$BMI>=25 & data2$BMI<30 ] = 3
# Obese
data2$BMICategory[ data2$BMI>=30 ] = 4
```

Next we can apply labels and convert the column to a factor, making it easier to read. The levels and labels parameters are telling our factor to associate the values of 1-4 with the labels we made earlier.

```{r}
data2$BMICategory = factor(data2$BMICategory,
                           levels = c(1,2,3,4),
                           labels = BMICategories)
```

We can see now our new column now.
```{r}
summary(data2$BMICategory)
```

***

## Making a Binary from Continuous

Perhaps we would like to do something even simpler. We can make a new binary column called `HealthyWeight`.
```{r}
# Create new column as FALSE
data2$HealthyWeight = FALSE

# Convert individuals to TRUE 
data2$HealthyWeight[ data2$BMI>=18.5 & data2$BMI<25 ] = TRUE

summary(data2$HealthyWeight)
```

Notice we could also use our new `BMICategory` column!
```{r}
# Set column as FALSE
data2$HealthyWeight = FALSE

# Convert individuals to TRUE 
data2$HealthyWeight[ data2$BMICategory=="Normal Weight" ] = TRUE

summary(data2$HealthyWeight)
```

***

