---
title: "Occupational Prestige Data Description"
output:
  html_document:
    theme: united
    toc: true
    toc_float: true
css: laurens_styles.css
---

```{r, child="_Global-Options.Rmd"}
```
***
***
### Occupational Prestige Data
```{r, eval=FALSE}
tempData <- read.table(url("https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/prestigeData.txt"), header = TRUE)

summary(tempData)
```
#### Survey Variables
1.  **YEAR**: Year for this respondent\
2.  **ID**: Unique ID\
3.  **WEIGHT**: Combined survey weight variables: WTSSCOMP*OVERSAMP\

#### Respondent Variables
1.  **VHAPPY**: Respondent’s general happiness
    **1:** Very Happy **0:** Pretty Happy or Not too happy\
2.  **COHORT**: Respondent's year of birth\
3.  **AGE**: Age of Respondent\
6.  **SEX**: Respondent's sex\
7.  **MentalHealth**: How does the respondent describe their own mental health\
    **1:** Excellent **2:** Very Good **3:** Good **4:** Fair **5:** Poor\
8.  **BMI**: Body Mass Index (Collected for adults)\
9.  **HasHypertension**: Has the respondent been diagnosed with high blood pressure?\
    **1:** Yes **0:** No\
10. **HasDiabetes**: Has the respondent been diagnosed with diabetes?\
    **1:** Yes **0:** No\
11. **Drinks5Day**: Over the past 12 months, did the respondent have 5 or more alcoholic drinks per day?\
    **1:** Yes **0:** No\
12. **HealthInsurance**: What type of health insurance, if any, does the respondent have?\
    **1:** Any Private **2:** Public Only **3:** Uninsured\
13. **HaveProvider**: Respondent has a usual healthcare provider\
    **1:** Yes **0:** No\
14. **NotAffordHealthCare**: In the past 12 months, respondent did not receive or delayed medical care because of cost?\
    **1:** Yes **0:** No\
15. **FamIncome_Continuous**: Family total income from all sources\
16. **FamIncome_PercentPoverty**: Family income divided by the applicable poverty line based on family size and composition\
17. **FamIncome_Categorical**: Family income as a percent of the poverty line organized as a categorical variable\
    **1:** Poor/Negative **2:** Near Poor **3:** Low Income **4:** Middle Income **5:** High Income\
18. **TotalHealthExpenditure**: Total health care expenditures ($): all sources (public insurance, private insurance, and out-of-pocket)\

