---
title: "Direct Marketing Dataset Description"
output:
  html_document:
    theme: united
    toc: true
    toc_float: true
css: laurens_styles.css
---

```{r, child="_Global-Options.Rmd"}

```
***
***
### Direct Marketing dataset
```{r}
tempData <- read.csv(url("https://laurencipriano.github.io/IveyBusinessStatistics/Datasets/SubscribeBank.csv"), header = TRUE)

summary(tempData)
```


### Summary of dataset

   The data is related with direct marketing campaigns of a Portuguese banking institution. 
   The marketing campaigns were based on phone calls. Often, more than one contact to the same client was required, in order to access if the product (bank term deposit) would be (or not) subscribed. 
   
   
### Data dictionary
 
```{r, echo=FALSE}
library(knitr)
   # Create the dataframe
variables <- data.frame(
  Variable = c("Dependent variable", "subscribe", 
               "Predictor variables",
                "age", "job", "marital", "education", "default", "balance", 
                "housing", "loan", "contact", "day", "month", "duration", 
                "campaign", "pdays", "previous", "poutcome"
              ),
  Description = c("", 
                  "Has the client subscribed a term deposit? (binary: 'yes', 'no')", 
                  "",
                  "Age of the client (numeric)", 
                  "Type of job (categorical: 'admin.', 'unknown', 'unemployed',       'management', 'housemaid', 'entrepreneur', 'student', 'blue-collar',                    'self-employed', 'retired', 'technician', 'services')", 
                  "Marital status (categorical: 'married', 'divorced', 'single'; 'divorced' includes divorced or widowed)", 
                  "Education level (categorical: 'unknown', 'secondary', 'primary', 'tertiary')", 
                  "Has credit in default? (binary: 'yes', 'no')", 
                  "Average yearly balance, in euros (numeric)", 
                  "Has housing loan? (binary: 'yes', 'no')", 
                  "Has personal loan? (binary: 'yes', 'no')", 
                  "Contact communication type (categorical: 'unknown', 'telephone', 'cellular')", 
                  "Last contact day of the month (numeric)", 
                  "Last contact month of year (categorical: 'jan', 'feb', ..., 'nov', 'dec')", 
                  "Last contact duration, in seconds (numeric)", 
                  "Number of contacts during this campaign (numeric, includes last contact)", 
                  "Number of days since last contact from a previous campaign (numeric, -1 if not previously contacted)", 
                  "Number of contacts before this campaign (numeric)", 
                  "Outcome of the previous marketing campaign (categorical: 'unknown', 'other', 'failure', 'success')" 
                  )
)

# Render the table in R Markdown
kable(variables, caption = "List of Variables and Definitions")
```


   
### Dataset access details

This dataset is public available for research. The details are described in [Moro et al., 2011]. 
  Please include this citation if you plan to use this database:

  [Moro et al., 2011] S. Moro, R. Laureano and P. Cortez. Using Data Mining for Bank Direct Marketing: An Application of the CRISP-DM Methodology. 
  In P. Novais et al. (Eds.), Proceedings of the European Simulation and Modelling Conference - ESM'2011, pp. 117-121, Guimarães, Portugal, October, 2011. EUROSIS.

  Available at: [pdf] http://hdl.handle.net/1822/14838
                [bib] http://www3.dsi.uminho.pt/pcortez/bib/2011-esm-1.txt


1. Title: Bank Marketing

2. Sources
   Created by: Paulo Cortez (Univ. Minho) and Sérgio Moro (ISCTE-IUL) @ 2012
   
3. Past Usage:

  The full dataset was described and analyzed in:

  S. Moro, R. Laureano and P. Cortez. Using Data Mining for Bank Direct Marketing: An Application of the CRISP-DM Methodology. 
  In P. Novais et al. (Eds.), Proceedings of the European Simulation and Modelling Conference - ESM'2011, pp. 117-121, Guimarães, 
  Portugal, October, 2011. EUROSIS.

4. Relevant Information:

   The data is related with direct marketing campaigns of a Portuguese banking institution. 
   The marketing campaigns were based on phone calls. Often, more than one contact to the same client was required, in order to access if the product (bank term deposit) would be (or not) subscribed. 

   There are two datasets: 
      1) bank-full.csv with all examples, ordered by date (from May 2008 to November 2010).
      2) bank.csv with 10% of the examples (4521), randomly selected from bank-full.csv.
   The smallest dataset is provided to test more computationally demanding machine learning algorithms (e.g. SVM).

   The classification goal is to predict if the client will subscribe a term deposit (variable y).

5. Number of Instances: 45211 for bank-full.csv (4521 for bank.csv)
